Comment: Reasoning about Interference Between Units

### 2014-10-17 paper/introduction.tex
More work on drafting the piece

BFP recommends choosing ``effect increasing" test statistics ``that will be
small when the treated and control distributions in the adjusted data ... are
similar, and large when the distributions diverge." However, test statistics
that only compare distributional features of hypothesized potential outcomes
between individuals assigned to treatment and those assigned to control fail
to fully exploit the data at hand. The problem is that these test statistics
are  not necessarily increasing in spillover effects.

I demonstrate two results, each showing how effect increasing test statistics,
as defined by BFP, may provide little to no information about the spillover
parameter $\tau_2$.  Let us first consider the case when the test statistic is
the absolute mean difference between hypothesized uniformity trial outcomes
for treated and control individuals: $\T(\widetilde y_{ij}(\0,\Z), \Z') =
\left\vert \mu_1 ( \widetilde y_{ij}(\0,\Z), \Z' ) - \mu_0 ( \widetilde
  y_{ij}(\0,\Z), \Z' ) \right\vert.  $ Fix a realization of $\Z = \z$, and
further assume noncollinearity: $ \z \notin \{\z^I, (1-\z^I)\}$. We can see
that, for any hypothesized $\widetilde\tau_2$, there always exists a
$\widetilde\tau_1$ such that the $p$-value is 1. Expanding the test statistic,
$\T(\widetilde y_{ij}(\0,\z), \z) = \left\vert \mu_1 (y_{ij}(\z) -
  \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) -
  \widetilde\tau_2 z_{i(-j)} , \z) \right\vert = \left\vert \mu_1 (y_{ij}(\z)
  - \widetilde\tau_2 z_{i(-j)}, \z)- \mu_0 (\widetilde y_{ij}(\z) -
  \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 \right\vert.  $ Thus,
regardless of the value of $\widetilde\tau_2$, we can always find a
$\widetilde\tau_1$ that sets $\T(\widetilde y_{ij}(\0,\z), \z) = 0$. Given
that $\T(\widetilde y_{ij}(\0,\z), \Z')$ can only take values in
$\mathbb{R}^+$, then $p=1$, as any other possible randomization $\Z'$ will
yield a value of $\T(\widetilde y_{ij}(\0,\z), \Z') \geq 0$. It follows that,
given this test, we cannot rule out any value of $\tau_2$ as being
implausible.\footnote{In fact, this result does not depend on example in the
  setting, and holds for any randomization scheme coupled with a linear causal
  model of the form \eqref{causal}.}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=.9\textwidth]{twoDplots.pdf}
\caption{Proportion of $p$-values less than .05 for tests of joint hypotheses
about $\tau$ and $\beta$. Darker values mean rare rejection. White means
rejection always. Truth is shown at the intersection of the straight
lines  $\tau=.5, \beta=2$. Each panel shows a different test statistic. The
SSR Tests refer to eq, the KS tests refer to eq.  }\label{fig:ks}
\end{figure}

Test statistic selection is not trivial and BFP's contribution has effectively
opened a new line of research. Via counterexample, we can see that effect
increasing test statistics are not always appropriate for assessing hypotheses
about interference. BFP is correct in the assertion that, regardless of the
choice of test statistic selection, a set of implausible hypotheses is
identified by the procedure. But we should not be led to believe that, for any
given test statistic, that some hypotheses are more plausible than others.
Such inferences -- comparing hypotheses -- may depend on the test statistic
used, and not necessarily reflect the plausibility of the model at hand. This
point will gain importance as we now consider the limitations of sharp null
hypotheses.

>>>>

How should one choose a test statistic to assess relationships between complex
models of treatment propagation and data? General advice for test statistics
is that they should be powerful against relevant alternatives (find canonical
cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific advice about
large sample performance of certain classes of test statistics and BFP repeat
his general advice: `` Select a test statistic T that will be small when the
treated and control distributions in the adjusted data \ldots are
similar, and large when the distributions diverge.'' \citep[p.
107]{bowers2013sutva}. They should have noted, instead, that when models
involve more than one parameter, such as the models of treatment propagation across
networks, test statistics should be increasing in \emph{all} of the parameters
(for example, in both parameters capturing spillover and parameters capturing
direct experience with the treatment). Here, we also show a benefit from
taking into account the network structure in the test statistic as well as in
the model.

We demonstrate two results, each showing how effect increasing test statistics,
as defined by BFP, may provide little to no information about the spillover
parameter $\tau_2$.  Let us first consider the case when the test statistic is
the absolute mean difference between hypothesized uniformity trial outcomes
for treated and control individuals: $\T(\widetilde y_{ij}(\0,\Z), \Z') =
\left\vert \mu_1 ( \widetilde y_{ij}(\0,\Z), \Z' ) - \mu_0 ( \widetilde
  y_{ij}(\0,\Z), \Z' ) \right\vert.  $ Fix a realization of $\Z = \z$, and
further assume noncollinearity: $ \z \notin \{\z^I, (1-\z^I)\}$. We can see
that, for any hypothesized $\widetilde\tau_2$, there always exists a
$\widetilde\tau_1$ such that the $p$-value is 1. Expanding the test statistic,

\begin{flalign}
  \T(\widetilde y_{ij}(\0,\z), \z) = & \left\vert \mu_1 (y_{ij}(\z) - \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) -
  \widetilde\tau_2 z_{i(-j)} , \z) \right\vert \\
= & \left\vert \mu_1 (y_{ij}(\z)
  - \widetilde\tau_2 z_{i(-j)}, \z)- \mu_0 (\widetilde y_{ij}(\z) -
  \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 \right\vert.
\end{flalign}

Thus,
regardless of the value of $\widetilde\tau_2$, we can always find a
$\widetilde\tau_1$ that sets $\T(\widetilde y_{ij}(\0,\z), \z) = 0$. Given
that $\T(\widetilde y_{ij}(\0,\z), \Z')$ can only take values in
$\mathbb{R}^+$, then $p=1$, as any other possible randomization $\Z'$ will
yield a value of $\T(\widetilde y_{ij}(\0,\z), \Z') \geq 0$. It follows that,
given this test, we cannot rule out any value of $\tau_2$ as being
implausible.\footnote{In fact, this result does not depend on example in the
  setting, and holds for any randomization scheme coupled with a linear causal
  model of the form \eqref{causal}.}

@@

As an example of the performance of these new statistics, we re-analyze the
model and design from BFP. Recall that their model of treatment propagation
was:

\begin{equation}
\HH(\by_\bz, \bw, \beta, \tau) =
 \frac{\beta + (1 - w_i) (1 - \beta) \exp(- \tau^2 \bw^{T} \bS)}
      {\beta + (1 - z_i) (1 - \beta) \exp(- \tau^2 \bz^{T} \bS)} \by_\bz
\label{eq:spillovermodelA}
\end{equation}

Briefly, the model posits that treatment effects can depend on either direct
assignment to treatment ($\bz$) governed by $\beta$ or spillover as an
increasing (but flattening) function of the number of directly connected
treated neighbors ($\bz^{T} \bS$) and is governed by $\tau$. So, we have a
model with two parameters. The network used by BFP involves 256 nodes
connected in an undirected, random graph with node degree ranging from 0 to 10
(mean degree 4, 95\% of nodes with degree between 1 and 8, five nodes with
degree 0 [i.e. unconnected]).  Treatment is assigned to 50\% of the nodes at
completely at random in the BFP example.

The top row of figure~\ref{fig:twoD} compares the power of the SSR+Design test
statistic (upper left panel) to versions of this statistic that either only
include fixed node degree (SSR+Degree) , or no information about the network
at all (SSR). For each test statistic, we tested the hypothesis
$\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation test (i.e. we used
1000 permutations instead of all of them). We executed that test 10,000 times
for each pair of parameters.  The proportion of $p$-values from that test less
than .05 is plotted in Figure~\ref{fig:twoD}: darker values are fewer
rejections, lighter values record more rejections.  All of these test
statistics are valid --- they reject the true null of $\tau=.5,\beta=2$ no
more than 5\% of the time at $\alpha=.05$ --- the plots are darkest in the
area where the two lines showing the true parameters intersect. All of the
plots have some power to reject non-true alternatives --- as we can see with
the large white areas in all of the plots. However, only when we add
information about the number of treated neighbors to the SSR+Degree
statistic, do we see high power against all alternatives in the plane.

\begin{figure}[h!]
\centering
\includegraphics[width=.99\textwidth]{twoDplots.pdf}
\caption{Proportion of $p$-values less than .05 for tests of joint hypotheses
about $\tau$ and $\beta$. Darker values mean rare rejection. White means
rejection always. Truth is shown at the intersection of the straight
lines  $\tau=.5, \beta=2$. Each panel shows a different test statistic. The
SSR Tests refer to eq, the KS tests refer to eq.  }\label{fig:twoD}
\end{figure}

The bottom row of Figure~\ref{fig:twoD} demonstrates the power of the KS test.
The bottom right hand panel shows the test used in BFP. Again all of the tests
are valid in the sense of rejecting the truth no more than 5\% of the time
when $\alpha=.05$ although all of these tests are conservative: the SSR based
tests rejected the truth roughly 4\% of the 10,000 simulations but the KS
tests rejected the truth roughly 2\% of the time.  The KS+Design and
KS+Degree panels show the power of applying the KS test to residuals from linear
models including network degree only (the +Degree version) or degree and also
the number of treated neighbors (a quantity that changes for each simulation
of the experiment). That is, where as the SSR panels used the sum of squared
residuals after accounting for network degree and/or number of treated
neighbors, the KS+Design and KS+Degree panels apply the KS test to the raw
residuals. These panels show (1) that inclusion of a quantity from the true
model (number of treated neighbors) is not enough to increase power against
all alternatives to the level shown by the SSR+Design test statistic and (2)
that the KS tests and the SSR tests have different patterns of power --- the
KS tests might be less powerful in general (more darker areas on the plots),
but our general sense from the analysis accords with the general sense about
the KS test in the scholarly literature --- that it is relatively low powered
(cites).

We cannot say here whether the SSR+Design test will provide the best power
against relevant alternatives for all possible models of treatment effect
propagation and designs. However, we hope that this research note both
improves the application of the BFP approach and raises new questions for
research.  Via counterexample, we can see that effect
increasing test statistics are not always appropriate for assessing hypotheses
about interference. BFP is correct in the assertion that, regardless of the
choice of test statistic selection, a set of implausible hypotheses is
identified by the procedure. But we should not be led to believe that, for any
given test statistic, that some hypotheses are more plausible than others.
Such inferences -- comparing hypotheses -- may depend on the test statistic
used, and not necessarily reflect the plausibility of the model at hand.
